---
title: "School Scores"
---

```{r}
library(tidyverse)
library(mosaic)
library(skimr)
library(ggformula)
library(GGally)
library(ggplot2)
```

## Importing Data

```{r}


score<- read.csv(file = '../../datasets/school-scores-data.csv')

(score)
```

```{r}
inspect(score)
glimpse(score)
skim(score)
```

## Data Dictionary

## -----------------------------------------------------------------

### Qualitative Variables:

#### State.Name (For IDENTIFYING the State of student)

#### State.Code (For IDENTIFYING the State by initials)

## -------

### Quantitative Variables: 

### Literally all other rows

## -----------------------------------------------------------------

## Observations

### It has a LOT of variables. Mostly quantitative variables that are scores of (along with number of) candidates in various tests & subjects. 

## ------------------------------------

## Research & Hypothesis

### Target Variables could be Total.Math, Total.Verbal because they give a summary of marks of students in that field. 

### Predictor Variables could be the Family.Income.More.Than.000 as they may influence test scores. Also variables such as Academic.Subjects.Average.GPA could be the predictors. (All subjects)

### They must have surveyed schools of every state to dig for student grades & their family background.

## ------------------------------------

# Table 1

## Code

```{r}

#| label: data-preprocessing

#CODE:

#FOR GETTING THE TABLE WITH SCORES BY SUBJECT


# Extracting the two columns

compact_score <- score[, c("Academic.Subjects.Arts.Music.Average.GPA", "Academic.Subjects.English.Average.GPA","Academic.Subjects.Foreign.Languages.Average.GPA","Academic.Subjects.Mathematics.Average.GPA","Academic.Subjects.Natural.Sciences.Average.GPA","Academic.Subjects.Social.Sciences.History.Average.GPA")]

# Renaming the columns because the names are annoyingly big

colnames(compact_score) <- c("Music", "English","Foreign Languages","Mathematics","NaturalSciences","SocialSciences")


head(compact_score)


```

## ----------------------------------------------------------------------------

# Chart 1

### It is like an Uncertainty Line Graph

## Questions:

### What is the correlation between students' GPA in a subject & the number of years for that subject?

## --------------------------------------------------------------------

# Table 2

## Code

#### Attempt 1

```{r}

#Again, extracting columns for info about Math scores & Family Incomes

#incomemath <- score[,c('State.Code','Family.Income.Less.than.20k.Math','Family.Income.Between.20.40k.Math','Family.Income.Between.40.60k.Math','Family.Income.Between.60.80k.Math','Family.Income.Between.80.100k.Math','Family.Income.More.than.100k.Math'	)]


#colnames(incomemath) <- c("Alabama","<20k","20-40k", "40-60k","60-80k","80-100k",">100k")


#incomemath %>%  filter(State.Code=='AL')
```

#### Attempt 2

```{r}
#library(dplyr)

# Assuming 'State.Code' is part of the 'score' dataframe
# Extract the math columns based on family income
#incomemath <- score[, c('Family.Income.Less.than.20k.Math',
                      #  'Family.Income.Between.20.40k.Math',
                       # 'Family.Income.Between.40.60k.Math',
                       # 'Family.Income.Between.60.80k.Math',
                      #  'Family.Income.Between.80.100k.Math',
                      #  'Family.Income.More.than.100k.Math')]

# Rename the columns for compactness
#colnames(incomemath) <- c("<20k", "20-40k", "40-60k", "60-80k", "80-100k", ">100k")

# Combine 'incomemath' with 'State.Code' for filtering purposes
# We assume that 'State.Code' exists in 'score'
#incomemath <- cbind(State.Code = score$State.Code, incomemath)

# Filter the row where 'State.Code' is 'AL'
#result <- incomemath %>% filter(State.Code == 'AL')

# Display the result as a table
#print(result)

```

#### Attempt 3

```{r}
library(dplyr)
library(tidyr)

# Extract math columns 
incomemath <- score[c('Family.Income.Less.than.20k.Math',
                        'Family.Income.Between.20.40k.Math',
                        'Family.Income.Between.40.60k.Math',
                        'Family.Income.Between.60.80k.Math',
                        'Family.Income.Between.80.100k.Math',
                        'Family.Income.More.than.100k.Math')]

# Rename the columns for compactness
colnames(incomemath) <- c("<20k", "20-40k", "40-60k", "60-80k", "80-100k", ">100k")



# Pivot to long format
long_format <- incomemath %>%
  pivot_longer(cols = everything(),      # Pivot all remaining columns
               names_to = "Income_Class", # Create a new column 'Income.Range' for the previous column names
               values_to = "Math_Scores")  # Store the values in a new column 'Math.Score'

# Display the result
head(long_format)

```

# Chart 2

### It is a Box-Plot

## ---------------------------

## Questions

### How is the distribution of students' marks in maths exam across all income levels? 

### Does the family income predict a student's success rate in mathematics? Or are the differences in their scores insignificant?

### Which income classes generally have highest and lowest math scores?

## ---------------------------------------

## Code

```{r}

 long_format %>% 
    gf_boxplot( reorder(Income_Class,Math_Scores) ~ Math_Scores ,
    fill = ~Income_Class,
    alpha = 0.5
  ) %>%

  gf_labs(
    title = "Math Scores vs Family Income",
   
  ) %>% 
 
  gf_labs(
    y = "Income Classes",
    x = "Scores in Math"
  )
```

### Here we have grouped number of students into boxes with income classes on Y Axis .

### (Also segregated by color.) And their Math Scores are on X Axis. (Collected across all states)

## ---------------------------

## Insights

### There are trends in math scores according to the income. Almost everywhere, the math scores are directly proportional to the family income. 

### Most of the students with 100k+ dollars at home have gotten \>500 Marks.

### More people from the 80-100k class have gotten the highest marks. (Except for the nerd in 20-40k range)

### Also, the scores become more clumped together & symmetrical as the income increases (boxes become thinner, along with the lines & median is in the centre)

### All groups of income have outliers, however the \<20k income group has the outliers clumped into one spot, while \>100k group has scattered outliers.

## ---------------------------

## Inference

### I had half-expected the income to directly influence the math scores, but not this visibly. Because not all rich kids are interested in studying.  I believe that it would be more intuitive if this graph was for college/grad students, as rich students (who like math) may have access to top colleges & hence resources to excel there. While a lot of poor students who like math may not have been able to afford college tuition. (Unless Scholarship) But that is also why the lower income categories have a wider distribution, so that only the very interested ones get good marks.

### Basically, the poorer you are, the more dedicated to math you have to be-compared to your classmates- in order to reach the top. \[The average student in the poor class gets poor marks.\] 

## \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
