---
title: "Cartoon Samples"
author: "Varad Akolkar"
---

```{r}
#| label: Setup
library(tidyverse) # Tidy data processing
library(ggformula) # Formula based plots
library(mosaic) # Data inspection and Statistical Inference
library(broom) # Tidy outputs from Statistical Analyses
library(infer) # Statistical Inference, Permutation/Bootstrap
library(patchwork) # Arranging Plots
library(ggprism) # Interesting Categorical Axes
library(supernova) # Beginner-Friendly ANOVA Tables
library(dplyr)
library(DescTools)
```

## Reading Data

```{r}
toon<- read_csv(file = '../../../datasets/cartoon.csv')
toon
glimpse(toon)
```

### We do not need to pivot this table since the cartoons are already in a single column.

```{r}
toon %>%  count(Cartoon)
toon %>%
    summarize(overall_mean = mean(Rating),count = n())
```

#### There is roughly 30 of each type which is enough data. Overall mean rating is 7.06.

## Data Analysis

```{r}

gf_histogram(~Rating,
  fill = ~Cartoon,
  data = toon, alpha = 1
) %>%  gf_vline(xintercept = ~ mean(Rating)) %>%
  
  
  gf_labs(
    title = "Histograms of Cartoon vs respective Rating. (0-10)",
    x = "Rating", y = "Count"
  ) %>%
  
  
  gf_text(16 ~ (mean(Rating)),
    label = "Overall Mean",
  ) %>%
  
  
  gf_refine(guides(fill = guide_legend(title = "Cartoon Name")))
```

### Looks like Doraemon ratings are all over the chart. Chota Bheem belongs in 6-8 range according to most people.

```{r}

gf_boxplot(
  data = toon,
  Rating ~ Cartoon,
  fill = ~Cartoon,
  alpha = 0.5
) %>% gf_vline(xintercept = ~ mean(Rating)) %>%
  
  gf_labs(
    title = "Boxplots of Rating Distributions vs Cartoons",
    x = "Cartoon", y = "Rating",
  ) %>%
  
  gf_refine(
    scale_x_discrete(guide = "prism_bracket"),
    guides(fill = guide_legend(title = "Cartoon"))
  )
```

## ANNOVA Test

```{r}

toon_anova <- aov(Rating ~ Cartoon, data = toon)

supernova::supernova(toon_anova)
```

```{r}

supernova::pairwise(toon_anova,
  correction = "Bonferroni", # Try "Tukey"
  alpha = 0.05, # 95% CI calculation
  var_equal = TRUE, # We'll see
  plot = TRUE
)
```

```{r}
# Calculate overall sum squares SST
toon_overall <- toon %>%
  summarise(
    overall_mean_rating = mean(Rating),
    # Overall mean across all readings
    # The Black Line

    SST = sum((Rating - overall_mean_rating)^2),
    n = n()
  ) 
toon_overall

SST <- toon_overall$SST
SST
```

```{r}
# Calculate sums of square errors *within* each group
# with respect to individual group means
toon_within_groups <- toon %>%
  group_by(Cartoon) %>%
  summarise(
    grouped_mean_rating = mean(Rating), 
    grouped_variance_rating = var(Rating),
    group_error_squares = sum((Rating - grouped_mean_rating)^2),
    n = n()
  )
toon_within_groups


toon_SSE <- toon_within_groups %>%
  summarise(SSE = sum(group_error_squares))
##
SSE <- toon_SSE$SSE
SSE
```

### The values here reflect what we saw in the histogram. D. Tales & Doraemon have a higher average rating than Chota Bheem. 

### Remember we said that Doraemon has its ratings all over the chart? That is why its variance is highest.

```{r}
SST
SSE
SSA <- SST - SSE
SSA
```

### Now we calculate degrees of freedom for all variances. The degrees of freedom is dependent on number of categories (3 Cartoons) & number of datapoints (30 for each).

```{r}
# Error Sum of Squares SSE
df_SSE <- toon %>%
  # Takes into account "unbalanced" situations
  # Where groups are not equal in size
  group_by(Cartoon) %>%
  summarise(per_group_df_SSE = n() - 1) %>%
  summarise(df_SSE = sum(per_group_df_SSE)) %>%
  as.numeric()


## Overall Sum of Squares SST
df_SST <- toon %>%
  summarise(df_SST = n() - 1) %>%
  as.integer()


# Treatment Sum of Squares SSA
k <- length(unique(toon$Cartoon))
df_SSA <- k - 1


##Degrees of Freedom:

df_SST
df_SSE
df_SSA
```

### The SST is one less than the total datapoints (91) as 1 degree of freedom goes into calculating the global mean.

### The SSE is sum of all categories with one less in each. 3\[(n1-1)+(n2-1)+(n3-1)\]

### The SSA is not the same as SSE-SST because we had an extra datapoint for Dragon Tales. Should have probably kept them equal. 

### Now we calculate the F-Statistic.

```{r}
# Combine the sum-square_error for each level of the factor
# Weighted by degrees of freedom **per level**

MSE <- toon_within_groups %>%
  summarise(mean_square_error = sum(group_error_squares / df_SSE)) %>%
  as.numeric()
MSE
```

```{r}
##
MSA <- SSA / df_SSA
MSA
```

```{r}

F_stat <- MSA / MSE
F_stat
```

#### Now we calculate the critical value of F-statistic.

```{r}
F_crit <-
  qf(
    p = (1 - 0.05 / 3), # Significance level is 5% + Bonferroni Correction
    df1 = df_SSA, # Numerator degrees of freedom
    df2 = df_SSE # Denominator degrees of freedom
  )
F_crit
F_stat
```

```{r}

mosaic::xpf(
  q = F_crit,
  df1 = df_SSA, df2 = df_SSE, method = "gg",
  log.p = FALSE, lower.tail = TRUE,
  return = "plot"
) %>%
  gf_vline(xintercept = F_crit) %>%
  gf_label(0.75 ~ 5,
    label = "F_critical",
    inherit = F, show.legend = F
  ) %>%
  gf_labs(
    title = "F distribution for Cartoons Data",
    subtitle = "F_critical = 4.290"
  )
```

### Since our F-Statistic is very low, & lower than F-Critical, we CANNOT state with confidence that Cartoon name has a direct effect on the rating.

### Which is obvious, since the rating is subjective by person, naturally a cartoon name does not guarantee a particular rating.

```{r}
supernova::equation(toon_anova)
```

## ANNOVA Assumptions

### ANOVA makes 3 fundamental assumptions:

#### 1. Data (and errors) are normally distributed.

#### 2. Variances are equal.

#### 3.Observations are independent.

#### We can check these using checks and graphs.

### Now we do the usual Shapiro Normality Test again to check the p.value.

```{r}
shapiro.test(x = toon$Rating)
```

### Since our p_value is much less than 0.05, we have to reject the NULL Hypothesis that the distribution is normal. 

### We can try doing the normality test at each level of factor.

```{r}
toon %>%
  group_by(Cartoon) %>%
  group_modify(~ .x %>%
    select(Rating) %>%
    as_vector() %>%
    shapiro.test() %>%
    broom::tidy())
```

### This time, the p_values for Doraemon & D.Tales are less than 0.05, so we have to reject the NULL Hypothesis of normality. (Even if there are supposedly enough samples.)

### We can check for residuals after model.

```{r}

toon_anova$residuals %>%
  as_tibble() %>%
  gf_dhistogram(~value, data = .) %>%
  gf_fitdistr()
##
toon_anova$residuals %>%
  as_tibble() %>%
  gf_qq(~value, data = .) %>%
  gf_qqstep() %>%
  gf_qqline()
##
shapiro.test(toon_anova$residuals)
```

### The residuals are not normal either. (P_value=0.0003\<0.05)

### We now check for similar Variance. To see if variances of each rating level across all cartoons are similar, we can use the Levenue Test or the Fligner-Killeen Test. (Since the data is not normally distributed)

### NULL Hypothesis: Data have similar variance

```{r}
toon %>%
  group_by(Cartoon) %>%
  summarise(variance = var(Rating))

DescTools::LeveneTest(Rating ~ Cartoon, data = toon)

fligner.test(Rating ~ Cartoon, data = toon)
```

### Looks like we have very different variances. As I said earlier, personal opinions.

## Effect Size

### We look for the actual effect sizes in Annova using our earlier graph.

```{r}

toon_supernova <-
  supernova::pairwise(toon_anova,
    plot = TRUE,
    alpha = 0.05,
    correction = "Bonferroni"
  )

toon_supernova
```

## ANNOVA Permutation Tests

### We use a permutation test to see the significance of effect size for all Cartoons & their normality.

#### We shuffle the cartoons randomly between the Ratings and repeat the ANOVA test each time and calculate the F-statistic. The Null distribution is the distribution of the F-statistic over the many permutations and the p-value is given by the proportion of times the F-statistic equals or exceeds that observed.

```{r}
observed_infer <-
  toon %>%
  specify(Rating ~ Cartoon) %>%
  hypothesise(null = "independence") %>%
  calculate(stat = "F")
observed_infer
```

### Now we use infer command to generate a NULL distribution using permutation of the Cartoon factor.

```{r}
null_dist_infer <- toon %>%
  specify(Rating ~ Cartoon) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 4999, type = "permute") %>%
  calculate(stat = "F")
##
null_dist_infer
```

```{r}
##
null_dist_infer %>%
  visualise(method = "simulation") +
  shade_p_value(obs_stat = observed_infer$stat, direction = "right") +
  scale_x_continuous(trans = "log10", expand = c(0, 0)) +
  coord_cartesian(xlim = c(0.2, 500), clip = "off") +
  annotation_logticks(outside = FALSE) 
```

### As seen, the infer based permutation test also shows that the permutationally generated F-statistics are close to which was observed. The effect of Cartoon barely registers.

### Hence, we CANNOT confidently say that the cartoon name has direct effect on its rating. The NULL Hypothesis prevails!

## ----------------------------------------------------------------------------
