---
title: "Inference"
author: "Varad"
---

```{r}
#Setup
library(tidyverse)
library(mosaic)
library(ggformula)
library(infer)
library(broom) # Clean test results in tibble form
library(resampledata) # Datasets from Chihara and Hesterberg's book
library(openintro) # More datasets
```

# Case Study 1: Toy Data

```{r}
set.seed(40)  # for replication
# Data as individual vectors ( for t.tests etc)
y <- rnorm(50, mean = 2, sd = 2)

# And as tibble too
mydata <- tibble(y = y)


x <- rnorm(500, mean = 2, sd = 2)

# And as tibble too
mydata2 <- tibble(x = x)

mydata
mydata2
```

```{r}

mydata %>%
    gf_density(~y) %>%
    gf_fitdistr(dist = "dnorm") %>%
    gf_labs(title = "Densities of Original Data Variables", subtitle = "Compared with Normal Density")

mydata2 %>%
    gf_density(~x) %>%
    gf_fitdistr(dist = "dnorm") %>%
    gf_labs(title = "Densities of 500 Original Data Variables", subtitle = "Compared with Normal Density")
```

### We see that upon increasing the sample size from 50 to 500, the distribution approaches a Bell Curve.

```{r}
# t-test
t1 <- mosaic::t_test(
          y, # Name of variable
          mu = 0, # belief of population mean
          alternative = "two.sided") %>% # Check both sides
  
  broom::tidy() # Make results presentable, and plottable!!
t1
```

# Case Study 2: Exam Data

```{r}
data("exam_grades")
exam_grades
```

```{r}

exam_grades %>%
    gf_density(~course_grade) %>%
    gf_fitdistr(dist = "dnorm") %>%
    gf_labs(title = "Density of Course Grade", subtitle = "Compared with Normal Density")
```

```{r}
stats::shapiro.test(x = exam_grades$course_grade) %>%
    broom::tidy()

```

```{r}
library(nortest)
# Especially when we have >= 5000 observations
nortest::ad.test(x = exam_grades$course_grade) %>%
    broom::tidy()
```

```{r}
# t-test
t4 <- mosaic::t_test(
          exam_grades$course_grade, # Name of variable
          mu = 80, # belief
          alternative = "two.sided") %>% # Check both sides
broom::tidy()
t4
```

# ---------------------------------------

```{r}
library(tidyverse) # Data Processing in R
library(mosaic) # Our workhorse for stats, sampling
library(skimr) # Good to Examine data
library(ggformula) # Formula interface for graphs

# load the NHANES data library
library(NHANES)
library(infer)
```

```{r}
data("NHANES")
glimpse(NHANES)
```

```{r}
NHANES_adult <-
  NHANES %>%
  distinct(ID, .keep_all = TRUE) %>%
  filter(Age >= 18) %>%
  select(Height) %>%
  drop_na(Height)
NHANES_adult
```

```{r}
# NHANES_adult is assumed population
pop_mean <- mosaic::mean(~Height, data = NHANES_adult)

pop_sd <- mosaic::sd(~Height, data = NHANES_adult)

pop_mean

pop_sd
```

```{r}


sample_50 <- mosaic::sample(NHANES_adult, size = 50) %>%
  select(Height)
sample_50
sample_mean_50 <- mean(~Height, data = sample_50)
sample_mean_50
# Plotting the histogram of this sample
sample_50 %>%
  gf_histogram(~Height, bins = 10) %>%
  gf_vline(
    xintercept = ~sample_mean_50,
    color = "purple"
  ) %>%
  gf_vline(
    xintercept = ~pop_mean,
    colour = "black"
  ) %>%
  gf_label(7 ~ (pop_mean + 8),
    label = "Population Mean",
    color = "black"
  ) %>%
  gf_label(7 ~ (sample_mean_50 - 8),
    label = "Sample Mean", color = "purple"
  ) %>%
  gf_labs(
    title = "Distribution and Mean of a Single Sample",
    subtitle = "Sample Size = 50"
  )
```

```{r}
sample_50_500 <- do(500) * {
  sample(NHANES_adult, size = 50) %>%
    select(Height) %>% # drop sampling related column "orig.id"
    summarise(
      sample_mean = mean(Height),
      sample_sd = sd(Height),
      sample_min = min(Height),
      sample_max = max(Height)
    )
}
sample_50_500
dim(sample_50_500)
```

```{r}


sample_50_500 %>%
  gf_point(.index ~ sample_mean,
    color = "purple",
    title = "Sample Means are close to the Population Mean",
    subtitle = "Sample Means are Random!",
    caption = "Grey lines represent our 500 samples"
  ) %>%
  gf_segment(
    .index + .index ~ sample_min + sample_max,
    color = "grey",
    linewidth = 0.3,
    alpha = 0.3,
    ylab = "Sample Index (1-500)",
    xlab = "Sample Means"
  ) %>%
  gf_vline(
    xintercept = ~pop_mean,
    color = "black"
  ) %>%
  gf_label(-25 ~ pop_mean,
    label = "Population Mean",
    color = "black"
  )
##

sample_50_500 %>%
  gf_point(.index ~ sample_sd,
    color = "purple",
    title = "Sample SDs are close to the Population Sd",
    subtitle = "Sample SDs are Random!",
  ) %>%
  gf_vline(
    xintercept = ~pop_sd,
    color = "black"
  ) %>%
  gf_label(-25 ~ pop_sd,
    label = "Population SD",
    color = "black"
  ) %>%
  gf_refine(lims(x = c(4, 16)))
```

```{r}

sample_50_500 %>%
  gf_dhistogram(~sample_mean, bins = 30, xlab = "Height") %>%
  gf_vline(
    xintercept = pop_mean,
    color = "blue"
  ) %>%
  gf_label(0.01 ~ pop_mean,
    label = "Population Mean",
    color = "blue"
  ) %>%
  gf_labs(
    title = "Sampling Mean Distribution",
    subtitle = "500 means"
  )


# How does this **distribution of sample-means** compare with the
# overall distribution of the population?
#
sample_50_500 %>%
  gf_dhistogram(~sample_mean, bins = 30, xlab = "Height") %>%
  gf_vline(
    xintercept = pop_mean,
    color = "blue"
  ) %>%
  gf_label(0.01 ~ pop_mean,
    label = "Population Mean",
    color = "blue"
  ) %>%
  ## Add the population histogram
  gf_histogram(~Height,
    data = NHANES_adult,
    alpha = 0.2, fill = "blue",
    bins = 30
  ) %>%
  gf_label(0.025 ~ (pop_mean + 20),
    label = "Population Distribution", color = "blue"
  ) %>%
  gf_labs(title = "Sampling Mean Distribution", subtitle = "Original Population overlay")
```

```{r}
# set.seed(12345)

samples_08_1000 <- do(1000) * mean(resample(NHANES_adult$Height, size = 08))

samples_16_1000 <- do(1000) * mean(resample(NHANES_adult$Height, size = 16))

samples_32_1000 <- do(1000) * mean(resample(NHANES_adult$Height, size = 32))

samples_64_1000 <- do(1000) * mean(resample(NHANES_adult$Height, size = 64))

# samples_128_1000 <- do(1000) * mean(resample(NHANES_adult$Height, size = 128))

# Quick Check
head(samples_08_1000)
```

```{r}


# Let us overlay their individual histograms to compare them:
p5 <- gf_dhistogram(~mean,
  data = samples_08_1000,
  color = "grey",
  fill = "dodgerblue", title = "N = 8"
) %>%
  gf_fitdistr(linewidth = 1) %>%
  gf_vline(
    xintercept = pop_mean, inherit = FALSE,
    color = "blue"
  ) %>%
  gf_label(-0.025 ~ pop_mean,
    label = "Population Mean",
    color = "blue"
  ) %>%
  gf_theme(scale_y_continuous(expand = expansion(mult = c(0.08, 0.02))))
##
p6 <- gf_dhistogram(~mean,
  data = samples_16_1000,
  color = "grey",
  fill = "sienna", title = "N = 16"
) %>%
  gf_fitdistr(linewidth = 1) %>%
  gf_vline(
    xintercept = pop_mean,
    color = "blue"
  ) %>%
  gf_label(-.025 ~ pop_mean,
    label = "Population Mean",
    color = "blue"
  ) %>%
  gf_theme(scale_y_continuous(expand = expansion(mult = c(0.08, 0.02))))
##
p7 <- gf_dhistogram(~mean,
  data = samples_32_1000,
  na.rm = TRUE,
  color = "grey",
  fill = "palegreen", title = "N = 32"
) %>%
  gf_fitdistr(linewidth = 1) %>%
  gf_vline(
    xintercept = pop_mean,
    color = "blue"
  ) %>%
  gf_label(-.025 ~ pop_mean,
    label = "Population Mean", color = "blue"
  ) %>%
  gf_theme(scale_y_continuous(expand = expansion(mult = c(0.08, 0.02))))

p8 <- gf_dhistogram(~mean,
  data = samples_64_1000,
  na.rm = TRUE,
  color = "grey",
  fill = "violetred", title = "N = 64"
) %>%
  gf_fitdistr(linewidth = 1) %>%
  gf_vline(
    xintercept = pop_mean,
    color = "blue"
  ) %>%
  gf_label(-.025 ~ pop_mean,
    label = "Population Mean", color = "blue"
  ) %>%
  gf_theme(scale_y_continuous(expand = expansion(mult = c(0.08, 0.02))))

# patchwork::wrap_plots(p5,p6,p7,p8)
p5
p6
p7
p8
```

```{r}
mean(~mean, data = samples_08_1000) # Mean of means!!!;-0
mean(~mean, data = samples_16_1000)
mean(~mean, data = samples_32_1000)
mean(~mean, data = samples_64_1000)
pop_mean
```

```{r}
pop_sd
sd(~mean, data = samples_08_1000)
sd(~mean, data = samples_16_1000)
sd(~mean, data = samples_32_1000)
sd(~mean, data = samples_64_1000)
```

```{r}
pop_sd
pop_sd / sqrt(8)
pop_sd / sqrt(16)
pop_sd / sqrt(32)
pop_sd / sqrt(64)
```

```{r}
sample_08 <- mosaic::sample(NHANES_adult, size = 8) %>%
  select(Height)
sample_16 <- mosaic::sample(NHANES_adult, size = 16) %>%
  select(Height)
sample_32 <- mosaic::sample(NHANES_adult, size = 32) %>%
  select(Height)
sample_64 <- mosaic::sample(NHANES_adult, size = 64) %>%
  select(Height)
##
sd(~Height, data = sample_08)
```

```{r}
sd(~Height, data = sample_16)
sd(~Height, data = sample_32)
sd(~Height, data = sample_64)
```

```{r}
pop_sd <- sd(~Height, data = NHANES_adult)
pop_sd
sd(~Height, data = sample_08) / sqrt(8)
sd(~Height, data = sample_16) / sqrt(16)
sd(~Height, data = sample_32) / sqrt(32)
sd(~Height, data = sample_64) / sqrt(64)
```

```{r}


tbl_1 <- get_ci(samples_08_1000, level = 0.95)
tbl_2 <- get_ci(samples_16_1000, level = 0.95)
tbl_3 <- get_ci(samples_32_1000, level = 0.95)
tbl_4 <- get_ci(samples_64_1000, level = 0.95)
rbind(tbl_1, tbl_2, tbl_3, tbl_4) %>%
  rownames_to_column("index") %>%
  cbind("sample_size" = c(8, 16, 32, 64)) %>%
  gf_segment(index + index ~ lower_ci + upper_ci) %>%
  gf_vline(xintercept = pop_mean) %>%
  gf_labs(
    title = "95% Confidence Intervals for the Mean",
    subtitle = "Varying samples sizes 8-16-32-64",
    y = "Sample Size", x = "Mean Ranges"
  ) %>%
  gf_refine(scale_y_discrete(labels = c(8, 16, 32, 64))) %>%
  gf_refine(annotate(geom = "label", x = pop_mean + 1.75, y = 1.5, label = "Population Mean"))
```

```{r}

sample_mean <- mean(~Height, data = sample_16)
se <- sd(~Height, data = sample_16) / sqrt(16)
#
xqnorm(
  p = c(0.025, 0.975),
  mean = sample_mean,
  sd = sd(~Height, data = sample_16),
  return = c("plot"), verbose = F
) %>%
  gf_vline(xintercept = ~pop_mean, colour = "black") %>%
  gf_vline(xintercept = mean(~Height, data = sample_16), colour = "purple") %>%
  gf_labs(title = "Confidence Intervals and the Bell Curve. N=16") %>%
  gf_refine(
    annotate(geom = "label", x = pop_mean + 15, y = 0.05, label = "Population Mean"),
    annotate(geom = "label", x = sample_mean - 15, y = 0.05, label = "Sample Mean", colour = "purple")
  )
```

```{r}
pop_mean
se <- sd(~Height, data = sample_16) / sqrt(16)
mean(~Height, data = sample_16) - 2.0 * se
mean(~Height, data = sample_16) + 2.0 * se
```

## **Workflow**

Thus if we want to estimate a *population mean*:

-   we take one **random** sample from the population

-   we calculate the estimate from the sample

-   we calculate the sample-sd

-   we calculate the *Standard Error* as sample−sdn

-   we calculate 95% confidence intervals for the *population parameter* based on the formula CI95%=sample.mean±2∗SE.

-   Since Standard Error decreases with sample size, we need to make our sample of adequate size.( n=30 seems appropriate in most cases. Why?)

-   And we do not have to worry about the distribution of the population. It need not be normal !!

## ---------------------------------------------------------
